Risk Perceptions for Wearable Devices

### 2015-03-06 tex-inputs/conclusion.tex
finish incorporating comments

Our 1,784 Internet users for their perceptions of wearable devices was the first large-scale study to investigate user security and privacy concerns for wearable devices. We identify that participants are most concerned about protecting their financial information, photos, and videos and least concerned about demographic or biometric data. We find that participants have a significant difference in perception of risk when data is being shared with an application's server versus other human recipients, and also that certain data is more comfortably shared with certain recipients. Participants' self-reported concerns of categories and perceptions of new wearable device capabilities are also presented to give a sense of the relevant user concerns, which can be used to guide future research in privacy with respect to wearable devices, especially warnings, notifications, and permissions. We perform two regression model analyses with respect to the type of data and the associated risks.

>>>>

Our survey of 1,784 Internet users for their perceptions of wearable devices was the first large-scale study to investigate user security and privacy concerns for wearable devices. We identify that participants are most concerned about protecting their financial information, photos, or videos and least concerned about demographic or biometric data. We find that participants have a significant difference in perception of risk when data is being shared with an application's server versus other human recipients, and also that certain data is more comfortably shared with certain recipients. Participants' self-reported concerns of new wearable device capabilities are also presented to give a sense of the relevant user concerns, which can be used to guide future research in privacy with respect to wearable devices, especially warnings, notifications, and permissions. We perform two regression model analyses with respect to the type of data and the associated risks to verify our results.

### 2015-03-06 tex-inputs/discussion.tex
finish incorporating comments

\section{Discussion}
Here, we discuss complementary future research directions in fields of privacy, ubiquitous computing, and user studies, along with specific limitations of this survey.

\subsection{Interpreting the Survey Results}
One of the main limitations of this work is that our participants might not have interest in or knowledge of wearables and their capabilities. 83\% of our participants reported that they do not own a wearable device. These participants may have underestimated or overestimated the risk perceived in various scenarios. People may be overreacting to recent events for scenarios \footnote{During our study, there were many stories covering injuries from exploding batteries (http://www.bloomberg.com/news/articles/2014-08-11/exploding-lithium-batteries-riskier-to-planes-research), which were explicitly and repeated mentioned when self-reporting concerns.}. People may be underestimating the risk of sharing certain data due to unawareness of what can be inferred from the data, or not have an idea of how to rate a new technology with respect to familiar ones. Biometrics were generally not a concern for our participants, although there are many security and privacy implications \cite{prabhakar2003biometric}.

We believed that getting a representative survey base was a useful endeavor. We could have easily recruited only wearables owners or people specifically interested in wearables. However, that will also have its own biases and limitations, since this does not reflect the general population. At the time of this writing, about 85\% of the general population do not own wearable devices \cite{Nilsen,WearableStatNews}, so our study reflective of the status quo. We expect user perceptions to change as rapidly as wearable technologies and the rate of adoption change.

Privacy concerns asked out of context differ from how users may react to these same concerns in real life \cite{norberg2007privacy, jensen2005privacy}. This is an unavoidable, yet important consideration of any study of this nature. This privacy paradox means that our findings may not be exactly representative of how upset users may be in real life, but do reflect their perceptions of wearable devices and various associated scenarios.

\subsection{Future Research Directions}
Further work can be done to expand various aspects of this study. Investigating more fine-grained data types (e.g. investigating if various types of location data, versus just location data in general) would be a useful endeavor to gain further insight into user perception. Adding more recipients, like ``advertisers'' or ``acquaintances'' may lead to more contrasting results.

@@

Wearables are still in infancy. Perceptions of situations and capabilities will change rapidly with advancements and increased exposure. However, videos and textual information are considered to be significantly sensitive by our participants, along with past participants of smartphone user perception studies. Various systems which detect and take actions for sensitive objects in photos and videos will be critical as wearables and other devices become more ubiquitous.

>>>>

\section{Discussion}
We discuss considerations when interpreting our results, the limitations of our work, and future work we hope to inspire.

\subsection{Interpreting the Survey Results}
One of the main limitations of this work is that our participants might not have interest in or knowledge of wearables and their capabilities. 83\% of our participants reported that they do not own a wearable device. These participants may have underestimated or overestimated the risk perceived in various scenarios. People may be overreacting to recent events for scenarios.\footnote{During our study, there were many stories covering injuries from exploding batteries (http://www.bloomberg.com/news/articles/2014-08-11/exploding-lithium-batteries-riskier-to-planes-research), which were explicitly mentioned in self-reported concerns.}. People may be underestimating the risk of sharing certain data due to unawareness of what can be inferred from the data, or not have an idea of how to rate a new technology with respect to familiar ones. Biometrics were generally not a concern for our participants, although there are many security and privacy implications \cite{prabhakar2003biometric}

We believed that getting a representative survey base was a useful endeavor. We could have easily recruited only wearables owners or people specifically interested in wearables. However, that will have its own biases and limitations, since this does not reflect the general population. At the time of this writing, about 85\% of the general population do not own wearable devices \cite{Nilsen,WearableStatNews}, so our study is reflective of the current population. We expect user perceptions to change as rapidly as wearable technologies and the rate of adoption change.

Privacy concerns asked out of context differ from how users may react to these same concerns in real life \cite{norberg2007privacy, jensen2005privacy}. This privacy paradox means that our findings may not be representative of how upset users may react to these scenarios in real life, but do reflect actual perceptions of wearable devices and various associated scenarios. This is an unavoidable, yet important distinction to make with of studies of our nature.

\subsection{Future Research Directions}
Further work can be done to expand various aspects of this study. Investigating more fine-grained data types (e.g. investigating various types of location data, versus just location data in general) would be a useful endeavor to gain further insight into user perception. Adding more recipients, like ``advertisers'' or ``acquaintances'' may lead to more contrasting results.

@@

Wearables are still in their infancy. Perceptions of situations and capabilities will change rapidly with advancements and increased exposure. However, videos and textual information are considered to be significantly sensitive by our participants, along with past participants of smartphone user perception studies. Various systems which detect and take actions for sensitive objects in photos and videos will be critical as wearables and other devices become more ubiquitous.

### 2015-03-06 tex-inputs/results-demographics.tex
finish incorporating comments

Participants' responses were correlated with demographic factors. We observed that the biggest predictor of participants' decisions to rate a scenario as very upsetting was their self-reported level of general privacy concerns, as determined by the IUIPC scale~\cite{malhotra2004internet}. A Spearman correlation yielded a statistically significant effect between average IUIPC scores with the VUR ($\rho=0.446$, $p<0.0005$), which suggests that people's responses to questions were mostly based on their privacy preferences. Additionally, we observed that age was a significant predictor of VUR ($\rho=0.121$, $p<0.0005$). We suspect that the effect of age is due to the significant correlation between age and IUIPC scores ($\rho=0.188$, $p<0.0005$); others have observed that older individuals tend to be more protective of their privacy~\cite{varian2005demographics}.

>>>>

Participants' responses were correlated with demographic factors. We observed that the biggest predictor of participants' decisions to rate a scenario as very upsetting was their self-reported level of general privacy concerns, as determined by the IUIPC scale~\cite{malhotra2004internet}. A Spearman correlation yielded a statistically significant effect between average IUIPC scores with the VUR ($\rho=0.446$, $p<0.0005$), which suggests that responses to questions were mostly based on privacy preferences. Additionally, we observed that age was a significant predictor of VUR ($\rho=0.121$, $p<0.0005$). We suspect that the effect of age is due to the significant correlation between age and IUIPC scores ($\rho=0.188$, $p<0.0005$); others have observed that older individuals tend to be more protective of their privacy~\cite{varian2005demographics}.

### 2015-03-06 tex-inputs/results-openended.tex
finish incorporating comments

This question was asked along with demographics questions (but before any IUIPC questions to avoid biasing). The participants were presented with a blank box to write in, with no character limit to their open-ended responses.

@@

In addition to privacy in a general sense, significant concerns included being unaware of what the device is collecting, doing, or which information it is using (Being Unaware), long-term health effects caused from wearing the device such as cancer from emf waves (Health), safety hazards from wearing the device, such as distractions which cause car accidents (Safety), resulting changes in social behaviors, such as dependencies on devices or spending less time with loved ones (Social Impact), the high financial cost of buying, replacing, or caring for the device (Financial), and information compromise (Security).

>>>>

This question was asked along with demographics questions (but before any IUIPC questions, which asked a lot of direct privacy-related questions, to avoid biasing the recipients). The participants were presented with a blank box to write in, with no character limit to their open-ended responses.

@@

In addition to privacy in a general sense, significant concerns included being unaware of what the device is collecting, doing, or which information it is using (Being Unaware), long-term health effects caused from wearing the device such as cancer from EMF waves (Health), safety hazards from wearing the device, such as distractions which cause car accidents (Safety), resulting changes in social behaviors, such as dependencies on devices or spending less time with loved ones (Social Impact), the high financial cost of buying, replacing, or caring for the device (Financial), and information compromise (Security).

### 2015-03-06 tex-inputs/results-techrank.tex
Use \columnwidth for figure width
finish incorporating comments

\subsection{Risk and Benefit Rankings} 
\begin{figure}[t]
	\centering
	\includegraphics[width=0.47\textwidth]{images/riskbenefit.pdf}
	\caption{Participants' median risk-benefit ratings of technologies examined by Fischhoff \etal\cite{Fischhoff}, which we used for calibration, alongside familiar technologies (e.g., laptops, the Internet, etc.), wearable technologies, as well as two specific wearable devices (Google Glass and the Cubetastic3000).}
	\label{fig:techplot}
\end{figure}

@@

Of the wearable technologies, privacy-invasive ones were perceived to be risky. Top risky technologies include facial recognition, the Internet, and discrete cameras, whereas the remainder of the technologies were seen as having minimal,  equivalent risk levels (i.e., a median of ``10''). The differences in risk that found between the different wearable-related technologies, are not tested for statistical significance, but given their minimal spread compared to the calibration options, the differences are negligible. Interestingly, privacy risks and are comparative to real physical risks; for instance, the capacity for facial detection on a wearable device is perceived to be almost as risky as interacting with a lawnmower.

People prompted to rate with respect to all considerations (see ~\ref{sec:prompt}), including physical harm risk to bystanders, financial cost, distress, misuse, or impact on public, personal, and private life, people may have still evaluated the risks with an emphasis toward physical risk and without an emphasis on privacy risk. Among the 5 presented options, the wearable-related one is the only one without some physical risk scenario, and physical risk is a clear, tangible risk to the users.

>>>>

\subsection{Risk and Benefit Rankings} 
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{images/riskbenefit.pdf}
	\caption{Participants' median risk-benefit ratings of technologies examined by Fischhoff \etal\cite{Fischhoff}, which we used for calibration, alongside familiar technologies (e.g., laptops, the Internet, etc.), wearable technologies, as well as two specific wearable devices (Google Glass and the Cubetastic3000).}
	\label{fig:techplot}
\end{figure}

@@

Of the wearable technologies, privacy-invasive ones were perceived to be risky. Top risky technologies include facial recognition, the Internet, and discrete cameras, whereas the remainder of the technologies were seen as having minimal,  equivalent risk levels (i.e., a median of ``10''). The differences in risk between the different wearable-related technologies, are not tested for statistical significance, but given their minimal spread compared to the calibration options, the differences are negligible. Interestingly, privacy risks and are comparable physical risks; for instance, the capacity for facial detection on a wearable device is perceived to be almost as risky as interacting with a lawnmower.

People  were prompted to rate technologies with respect to all considerations (see ~\ref{sec:prompt}), including physical harm risk to bystanders, financial cost, distress, misuse, or impact on public, personal, and private life, people may have still evaluated the risks with an emphasis toward physical risk and without an emphasis on privacy risk. Among the five presented options, the wearable-related one is the only one without some physical risk scenario, and physical risk is a clear, tangible risk to the users.
